---
title: "Email Cleaning"
---

# Load libraries ---------------------------------------------------------------

```{r}

library(tidyverse)

```

# Load data --------------------------------------------------------------------

```{r}

emails <- read_csv("data/emails_raw.csv")

```

# Summary statistics -----------------------------------------------------------

```{r}

# How many distinct DOIs are there?
emails %>% distinct(doi) %>% count()

# How many distinct email addresses are there?
emails %>% distinct(email) %>% count()

```

# Validate email addresses -----------------------------------------------------

```{r}

# Function for validation - see https://www.r-bloggers.com/validating-email-adresses-in-r/
# This is quite basic and doesn't catch all!
# LimeSurvey has email validation built in, so not too important, but good to get an overview
isValidEmail <- function(x) {
	grepl("\\<[A-Z0-9._%+-]+@[A-Z0-9.-]+\\.[A-Z]{2,}\\>", as.character(x), ignore.case=TRUE)
}

# Final cleaned email addresses
emails_clean <- emails %>%
  # clean up strings - remove whitespace, convert to lowercase
  mutate(email = stringr::str_trim(email),
         email = stringr::str_squish(email),
         email = stringr::str_to_lower(email)) %>%
  mutate(is_valid = isValidEmail(email)) %>%
  filter(is_valid == T) %>%
  distinct(email)

# save to csv
write_csv(emails_clean, "data/emails_clean.csv")

```

# Other ------------------------------------------------------------------------

```{r}

# Some fun - what are most popular email hosts used?
as_tibble(str_split_fixed(read_csv("data/emails_clean.csv") %>% 
                            pull(email), "@", 2)) %>%
  select(V2) %>%
  rename("host" = "V2") %>%
  count(host) %>%
  top_n(30, n) %>%
  ggplot(aes(x = host, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()

```

# Building email lists ---------------------------------------------------------

The list of clean email addresses is uploaded to LimeSurvey, and survey tokens are generated.
Here, we use the tokens to manually create columns for invitation and optout URLs.
These URLs are then added to each email via mail merge in MS Word.
Finally, the list of participants is split into files with batches of 500 for sending.

```{r}

emails_limesurvey <- read_csv("data/emails_limesurvey.csv") %>%
  mutate(invitation_url = paste0("https://tigereye.informatik.uni-kiel.de/limesurvey/index.php/564879?token=", token, "&lang=en"),
         optout_url = paste0("https://tigereye.informatik.uni-kiel.de/limesurvey/index.php/optout/tokens/564879?langcode=en&token=", token))

emails_limesurvey %>%
  group_by(g = ceiling(row_number() / 500)) %>% 
  do(write_csv(., paste0("data/email_batches/batch_", .$g[1], '.csv')))

```

Reminder handling: we use the same process as above, but filter for those who have not
opted out, who have not yet completed the survey, and who were not manually flagged (e.g.
due to unhappy email responses from the first invitation batches)

```{r}

emails_limesurvey_reminder <- read_csv("data/emails_limesurvey_reminder.csv") %>%
  filter(emailstatus == "OK",
         completed == "N",
         usesleft == 1,
         flagged == 0) %>%
  mutate(invitation_url = paste0("https://tigereye.informatik.uni-kiel.de/limesurvey/index.php/564879?token=", token, "&lang=en"),
         optout_url = paste0("https://tigereye.informatik.uni-kiel.de/limesurvey/index.php/optout/tokens/564879?langcode=en&token=", token))

emails_limesurvey_reminder %>%
  group_by(g = ceiling(row_number() / 500)) %>% 
  do(write_csv(., paste0("data/email_batches_reminder/batch_", .$g[1], '.csv')))

```



