---
title: "Main Analysis"
---

# Load libraries ---------------------------------------------------------------

```{r}

library(tidyverse)
library(patchwork)

```

# Themes -----------------------------------------------------------------------

```{r}

t <- theme(text = element_text(size = 10),
           plot.title = element_text(size = 10, face = "bold"),
           axis.title.x = element_text(size = 10, margin = margin(10, 0, 0, 0)),
           axis.title.y = element_text(size = 10, margin = margin(0, 10, 0, 0)),
           axis.text.x = element_text(size = 8),
           axis.text.y = element_text(size = 8),
           plot.margin = margin(10, 10, 10, 10))

theme_set(theme_minimal() + t)

```

# Helpers ----------------------------------------------------------------------

```{r}

# Factorise likert levels for survey responses
likert_factorize <- function(d) {
  d %>%
    mutate(response = factor(response, 
                             levels = c("Strongly agree",
                                        "Agree",
                                        "Neutral",
                                        "Disagree",
                                        "Strongly disagree")))
}

# Make survey plots with likert scales
plot_likert <- function(data, codes, title = NULL) {
  
  plot_labels <- codes %>%
    filter(code %in% colnames(data)) %>%
    mutate(code = fct_reorder(code, colnames(data)),
           label = as.character(label)) %>%
    pull(label)
  
  plot <- data %>%
    pivot_longer(., 
                 cols = starts_with("A"), 
                 names_to = "code", 
                 values_to = "response") %>%
    likert_factorize() %>%
    inner_join(codes, by="code") %>%
    droplevels() %>%
    group_by(label, response) %>%
    summarize(n = n()) %>%
    mutate(freq = n / sum(n)) %>%
    ungroup() %>%
    ggplot() +
    geom_bar(aes(x = label, y = freq, fill = response), 
             color = "grey50", width = 0.8, size = 0.25, 
             stat = "identity", position = "stack") +
    coord_flip() +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 60),
                     limits = rev(plot_labels)) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = colorspace::diverging_hcl(n = 5, 
                                                         palette = "Blue-Red", 
                                                         alpha = 0.8)) +
    labs(x = "", y = "", fill = "", 
         title = str_c(title, " (N = ", nrow(data), ")")) +
    guides(fill = guide_legend(reverse = T)) +
    theme(axis.title.x = element_blank(),
          legend.position="bottom")
  
  return(plot)
  
}

```

# Load full response data ------------------------------------------------------

```{r}

responses <- read_csv("data/response_data_anonymised.csv")

```

# Demographics -----------------------------------------------------------------

## Countries

```{r}

country <- responses %>%
  select(country = starts_with("Country")) %>%
  mutate(country = case_when(
    country == "the United Kingdom of Great Britain and Northern Ireland" ~ "UK",
    country == "United States of America" ~ "USA",
    T ~ country
  )) %>%
  count(country) %>%
  top_n(30, n) %>%
  ggplot() +
  geom_col(aes(x = reorder(country, n), y = n), 
           fill = "grey75", color = "grey25", width = 0.8, size = 0.25) +
  labs(y = "Participants", x = "", title = "Countries (Top 30)") +
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 30)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_flip(ylim = c(0, 400)) +
  ggsave("outputs/figures/partials/demographics_country.png", width = 4, height = 6)

```

## Gender

```{r}

gender <- responses %>%
  select(gender = starts_with("Gender")) %>% 
  count(gender) %>%
  ggplot() +
  geom_col(aes(x = reorder(gender, n), y = n), 
           fill = "grey75", color = "grey25", width = 0.8, size = 0.25) +
  labs(y = "Participants", x = "", title = "Gender") +
  scale_y_continuous(expand = c(0, 0)) +
  coord_flip(ylim = c(0, 1250)) +
  ggsave("outputs/figures/partials/demographics_gender.png", width = 4, height = 2)

```

## Career Status

```{r}

career_status <- responses %>%
  select(career_status = starts_with("CareerStatus")) %>%
  count(career_status) %>%
  ggplot() +
  geom_col(aes(x = reorder(career_status, n), y = n), 
           fill = "grey75", color = "grey25", width = 0.8, size = 0.25) +
  labs(y = "Participants", x = "", title = "Career Status") +
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 30)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_flip(ylim = c(0, 400))  +
  ggsave("outputs/figures/partials/demographics_career_status.png", 
         width = 4, height = 3)

```

## Institution Type

```{r}

institution <- responses %>%
  select(institution_type = starts_with("Institution")) %>% 
  count(institution_type) %>%
  ggplot() +
  geom_col(aes(x = reorder(institution_type, n), y = n), 
           fill = "grey75", color = "grey25", width = 0.8, size = 0.25) +
  labs(y = "Participants", x = "", title = "Institution Types") +
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 35)) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip(ylim = c(0, 1000)) +
  ggsave("outputs/figures/partials/demographics_institution_type.png", width = 4, height = 3)

```

## Discipline

```{r}

discipline <- responses %>%
  pivot_longer(cols = starts_with("Discipline"), 
               names_to = "label", 
               values_to = "response") %>%
  inner_join(read_csv("data/coding/Discipline.csv"), by = "label") %>%
  mutate(response = case_when(
    response == "Yes" ~ T,
    T ~ F
  )) %>%
  group_by(discipline) %>%
  summarize(n = sum(response)) %>%
  ggplot() +
  geom_col(aes(x = reorder(discipline, n), y = n), 
           fill = "grey75", color = "grey25", width = 0.8, size = 0.25) +
  labs(y = "Participants", x = "", title = "Disciplines") +
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 35)) +
  scale_y_continuous(expand = c(0,0)) +
  coord_flip(ylim = c(0,350)) +
  ggsave("outputs/figures/partials/demographics_discipline.png", width = 4, height = 6)

```

## Composite Figure 

```{r}

discipline + 
  country + 
  (gender / institution / career_status + plot_layout(heights = c(0.6, 1, 1.5))) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/demographics.png", width = 12, height = 6)

```

# Publishing behaviour ---------------------------------------------------------

```{r}

# Number of articles published

pub_levels = c("0", "1-2", "3-5", "6-10", "11-20", "21+")

plot_n_pub <- function(d, type) {
  
  d %>%
    na.omit() %>%
    mutate(n_pub = factor(n_pub, levels = pub_levels)) %>%
    count(n_pub) %>%
    ggplot() +
    geom_col(aes(x = n_pub, y = n), 
             fill = "grey75", color = "grey25", width = 0.8, size = 0.25)  +
    labs(y = "Participants", x = "", 
         title = str_wrap(str_c(type, ": journal articles published in past 5 years"), 50)) +
    coord_flip()
  
}

n_pub_cor <- responses %>%
  select(n_pub = matches("NPublishedCor\\.")) %>%
  plot_n_pub(., "Corresponding authorships")

n_pub_co <- responses %>%
  select(n_pub = matches("NPublishedCo\\.")) %>%
  plot_n_pub(., "Co-authorships")

# Proportion of articles deposited as preprints

dep_levels = c("None of these articles were also deposited as preprints*",
               "Some of these articles were also deposited as preprints*",
               "All of these articles were also deposited as preprints*")

dep_labels = c("None of these articles were also deposited as preprints",
               "Some of these articles were also deposited as preprints",
               "All of these articles were also deposited as preprints")

plot_p_dep <- function(d, type) {
  d %>%
    na.omit() %>%
    mutate(p_dep = factor(p_dep, levels = dep_levels, labels = dep_labels)) %>%
    count(p_dep) %>%
    ggplot() +
    geom_col(aes(x = reorder(p_dep, -p_dep), y = n),
             fill = "grey75", color = "grey25", width = 0.8, size = 0.25)  +
    labs(y = "Participants", x = "", 
         title = str_wrap(str_c(type, ": proportion of articles deposited as preprints"), 50)) +
    scale_x_discrete(label = function(x) str_wrap(x, 40)) +
    coord_flip()
}

p_dep_cor <- responses %>%
  select(p_dep = matches("PDepositedCor\\.")) %>%
  plot_p_dep(., "Corresponding authorships")

p_dep_co <- responses %>%
  select(p_dep = matches("PDepositedCo\\.")) %>%
  plot_p_dep(., "Co-authorships")

# Patchwork
n_pub_cor + n_pub_co + p_dep_cor + p_dep_co +
  plot_annotation(tag_levels = 'A') +
  plot_layout(heights = c(1.75, 1), ncol = 2) +
  ggsave("outputs/figures/publishing_behaviour.png",
         width = 12,
         height = 6)
```

# Preprint motivations ---------------------------------------------------------

## Data

```{r}

deposited_codes <- read_csv("data/coding/deposited.csv", col_types = "ff")

# Corresponding authorships
cor_all_deposited <- responses %>% 
  select(A = starts_with("CorAllDeposited")) %>%
  na.omit()
cor_some_deposited <- responses %>% 
  select(A = starts_with("CorSomeDeposited")) %>%
  na.omit()
cor_deposited <- bind_rows(cor_all_deposited, cor_some_deposited)

# Co-authorships
co_all_deposited <- responses %>% 
  select(A = starts_with("CoAllDeposited")) %>%
  na.omit()
co_some_deposited <- responses %>% 
  select(A = starts_with("CoSomeDeposited")) %>%
  na.omit()
co_deposited <- bind_rows(co_all_deposited, co_some_deposited)

```

## Decision making

```{r}

# This refers to items A1, A4, A5, A2 and A3

# Corresponding authorships
p_cor_deposited_decision <- plot_likert(
  cor_deposited %>% select(A1, A4, A5, A2, A3), 
  deposited_codes,
  "Corresponding Authorships")

p_co_deposited_decision <- plot_likert(
  co_deposited %>% select(A1, A4, A5, A2, A3), 
  deposited_codes,
  "Co-authorships")

p_deposited_decision <- p_cor_deposited_decision + 
  (p_co_deposited_decision + theme(axis.text.y = element_blank())) +
  plot_layout(ncol = 2, guides = "collect") & theme(legend.position = "bottom")

p_deposited_decision +
  ggsave("outputs/figures/deposited_decision.png", width = 12, height = 5)

```

## Motivations

```{r}

# Items A6, A7, A8, A9, A10

# What factors motivate preprint deposition (A6, A7, A8, A9, A10)
p_cor_deposited_motivation <- plot_likert(
  cor_deposited %>% select(A6, A7, A8, A9, A10), 
  deposited_codes,
  "Corresponding Authorships")

p_co_deposited_motivation <- plot_likert(
  co_deposited %>% select(A6, A7, A8, A9, A10), 
  deposited_codes,
  "Co-Authorships")

p_deposited_motivation <- p_cor_deposited_motivation + 
  (p_co_deposited_motivation + theme(axis.text.y = element_blank())) +
  plot_layout(ncol = 2, guides = "collect") & theme(legend.position = "bottom")

p_deposited_motivation +
  ggsave("outputs/figures/deposited_motivation.png", width = 12, height = 5)

```

## Benefits

```{r}

# Items A11, A12

p_cor_deposited_benefits <- plot_likert(
  cor_deposited %>% select(A11, A12), 
  deposited_codes,
  "Corresponding Authorships")

p_co_deposited_benefits <- plot_likert(
  co_deposited %>% select(A11, A12), 
  deposited_codes,
  "Co-Authorships")

p_deposited_benefits <- p_cor_deposited_benefits + 
  (p_co_deposited_benefits + theme(axis.text.y = element_blank())) +
  plot_layout(ncol = 2, guides = "collect") & theme(legend.position = "bottom")

p_deposited_benefits +
  ggsave("outputs/figures/deposited_benefits.png", width = 12, height = 3)

```

# Preprint demotivations -------------------------------------------------------

## Data

```{r}

not_deposited_codes <- read_csv("data/coding/not_deposited.csv", col_types = "ff")

# Corresponding authorships
cor_none_not_deposited <- responses %>% 
  select(A = starts_with("CorNoneNotDeposited")) %>%
  na.omit()
cor_some_not_deposited <- responses %>% 
  select(A = starts_with("CorSomeNotDeposited")) %>%
  na.omit()
cor_not_deposited <- bind_rows(cor_none_not_deposited, cor_some_not_deposited)

# Corresponding authorships
co_none_not_deposited <- responses %>% 
  select(A = starts_with("CoNoneNotDeposited")) %>%
  na.omit()
co_some_not_deposited <- responses %>% 
  select(A = starts_with("CoSomeNotDeposited")) %>%
  na.omit()
co_not_deposited <- bind_rows(co_none_not_deposited, co_some_not_deposited)

```

## Plots

```{r}

p_cor_not_deposited_demotivations <- plot_likert(
  cor_not_deposited,
  not_deposited_codes,
  "Corresponding Authorships")

p_co_not_deposited_demotivations <- plot_likert(
  co_not_deposited,
  not_deposited_codes,
  "Co-Authorships")

p_not_deposited_demotivations <- p_cor_not_deposited_demotivations + 
  (p_co_not_deposited_demotivations + theme(axis.text.y = element_blank())) +
  plot_layout(ncol = 2, guides = "collect") & theme(legend.position = "bottom")

p_not_deposited_demotivations +
  ggsave("outputs/figures/not_deposited_demotivations.png", width = 12, height = 8)


```

# Preprint comparisons ---------------------------------------------------------

## Data

```{r}

comparison_codes <- read_csv("data/coding/comparison.csv", col_types = "ff")

# Corresponding authorships
cor_comparison <- responses %>% 
  select(A = starts_with("CorSomeComparison")) %>%
  na.omit()

# Co-authorships
co_comparison <- responses %>% 
  select(A = starts_with("CoSomeComparison")) %>%
  na.omit()

```

## Plots

```{r}

p_cor_comparison <- plot_likert(
  cor_comparison,
  comparison_codes,
  "Corresponding Authorships")

p_co_comparison <- plot_likert(
  co_comparison,
  comparison_codes,
  "Co-Authorships")

p_comparison <- p_cor_comparison + 
  (p_co_comparison + theme(axis.text.y = element_blank())) +
  plot_layout(ncol = 2, guides = "collect") & theme(legend.position = "bottom")

p_not_deposited_demotivations +
  ggsave("outputs/figures/comparisons.png", width = 12, height = 6)

```

# Qualitative data -------------------------------------------------------------

Note: Raw datasets not provided as part of archived data due to potential for
participant identification

## Deposited

```{r}

cor_deposited_free <- read_csv("data/qualitative/DepositedFreeCoded.csv") %>%
  rename(
    "id" = starts_with("id."),
    "text" = starts_with("CorDepositedFree."),
    "code" = starts_with("CorDepositedFreeCode")) %>%
  filter(!is.na(code)) %>%
  select(id, text, code) %>%
  mutate(code = str_split(code, ";")) %>% 
  unnest(code) %>%
  mutate(code = str_trim(code))

n_cor_deposited_free <- n_distinct(cor_deposited_free$id)

cor_deposited_free %>%
  group_by(code) %>%
  summarise(n = n_distinct(id)) %>%
  ungroup() %>%
  mutate(prop = round(n * 100 / n_cor_deposited_free, 1))

co_deposited_free <- read_csv("data/qualitative/DepositedFreeCoded.csv") %>%
  rename(
    "id" = starts_with("id."),
    "text" = starts_with("CoDepositedFree."),
    "code" = starts_with("CoDepositedFreeCode")) %>%
  filter(!is.na(code)) %>%
  select(id, text, code) %>%
  mutate(code = str_split(code, ";")) %>% 
  unnest(code) %>%
  mutate(code = str_trim(code))

n_co_deposited_free <- n_distinct(co_deposited_free$id)

co_deposited_free %>%
  group_by(code) %>%
  summarise(n = n_distinct(id)) %>%
  ungroup() %>%
  mutate(prop = round(n * 100 / n_co_deposited_free, 1))

```

## Deposited

```{r}

cor_not_deposited_free <- read_csv("data/qualitative/NotDepositedFreeCoded.csv") %>%
  rename(
    "id" = starts_with("id."),
    "text" = starts_with("CorNotDepositedFree."),
    "code" = starts_with("CorNotDepositedFreeCode")) %>%
  filter(!is.na(code)) %>%
  select(id, text, code) %>%
  mutate(code = str_split(code, ";")) %>% 
  unnest(code) %>%
  mutate(code = str_trim(code))

n_cor_not_deposited_free <- n_distinct(cor_not_deposited_free$id)

cor_not_deposited_free %>%
  group_by(code) %>%
  summarise(n = n_distinct(id)) %>%
  ungroup() %>%
  mutate(prop = round(n * 100 / n_cor_not_deposited_free, 1))

co_not_deposited_free <- read_csv("data/qualitative/NotDepositedFreeCoded.csv") %>%
  rename(
    "id" = starts_with("id."),
    "text" = starts_with("CoNotDepositedFree."),
    "code" = starts_with("CoNotDepositedFreeCode")) %>%
  filter(!is.na(code)) %>%
  select(id, text, code) %>%
  mutate(code = str_split(code, ";")) %>% 
  unnest(code) %>%
  mutate(code = str_trim(code))

n_co_not_deposited_free <- n_distinct(co_not_deposited_free$id)

co_not_deposited_free %>%
  group_by(code) %>%
  summarise(n = n_distinct(id)) %>%
  ungroup() %>%
  mutate(prop = round(n * 100 / n_co_not_deposited_free, 1))

```

## Comparisons

```{r}

cor_comparison_free <- read_csv("data/qualitative/ComparisonFreeCoded.csv") %>%
  rename(
    "id" = starts_with("id."),
    "text" = starts_with("CorComparisonFree."),
    "code" = starts_with("CorComparisonFreeCode")) %>%
  filter(!is.na(code)) %>%
  select(id, text, code) %>%
  mutate(code = str_split(code, ";")) %>% 
  unnest(code) %>%
  mutate(code = str_trim(code))

n_cor_comparison_free <- n_distinct(cor_comparison_free$id)

cor_comparison_free %>%
  group_by(code) %>%
  summarise(n = n_distinct(id)) %>%
  ungroup() %>%
  mutate(prop = round(n * 100 / n_cor_comparison_free, 1))

co_comparison_free <- read_csv("data/qualitative/ComparisonFreeCoded.csv") %>%
  rename(
    "id" = starts_with("id."),
    "text" = starts_with("CoComparisonFree."),
    "code" = starts_with("CoComparisonFreeCode")) %>%
  filter(!is.na(code)) %>%
  select(id, text, code) %>%
  mutate(code = str_split(code, ";")) %>% 
  unnest(code) %>%
  mutate(code = str_trim(code))

n_co_comparison_free <- n_distinct(co_comparison_free$id)

co_comparison_free %>%
  group_by(code) %>%
  summarise(n = n_distinct(id)) %>%
  ungroup() %>%
  mutate(prop = round(n * 100 / n_co_comparison_free, 1))

```





